Grade Retention and Mixed Race

### 2022-08-18 products/paper/main.qmd
Add rought draft of role of class subsection

@@

All of these arguments relate to where we might expect the outcomes of biracial individuals to fall relative to their monoracial peers. However, the experience of hybridity itself may have positive and negative effects that do not tell us the location itself but rather may modify it in the context of other patterns. On the positive side, hybridity may enlarge the cultural "toolkit" (==CITATION==) of biracial individuals. This enlarged toolkit may allow such individuals to more easily navigate a variety of social situations (e.g. code switch). On the negative side, Stonequist's (==CITATION==) "marginal man" hypothesis suggests that the dissonance of an identity that crosses strong and salient racial boundaries will be experienced by biracial individuals as not fitting in anywhere. This may result in negative mental health experiences for biracial individuals that can spill over into other types of outcomes.

@@

-   In all of this, its crucially important to understand how other resources affect our understanding of the placement of biracial individuals. Biracial individuals are the products of interracial unions and individuals are not randomly selected into such unions. This can lead to underlying differences in the socioeconomic resources that biracial individuals possess relative to their monoracial peers. However, this selection can be quite complex. (DISCUSS CHENG AND LIVELY HERE?)

    -   Status exchange suggests that members of lower status racial groups will be positively selected into interracial unions, while higher status racial groups will be negatively selected. How this affects the relative resources of biracial individuals will depend then in some part on the relative strength of these phenomena, the gender balance in a given form of interracial unions, and how the gender of parents affects outcomes. Status exchange may also not be present in all forms of interracial unions.

@@

>>>>

## Where do Biracials Fit?

@@

All of these arguments relate to where we might expect the outcomes of biracial individuals to fall relative to their monoracial peers. However, the experience of hybridity itself may have positive and negative effects that do not tell us the location itself but rather may modify it in the context of other patterns. On the positive side, hybridity may enlarge the cultural "toolkit" (==CITATION==) of biracial individuals. This enlarged toolkit may allow such individuals to more easily navigate a variety of social situations (e.g. code switch). On the negative side, Stonequist's (==CITATION==) "marginal man" hypothesis suggests that the dissonance of an identity that crosses strong and salient racial boundaries will be experienced by biracial individuals as not fitting in anywhere. This liminality may result in negative mental health experiences for biracial individuals that may affect other types of outcomes as well.

@@

## The Role of Class

In addition to these specifically racial mechanisms that may drive the placement of biracial individuals, researchers must also pay careful attention to how socioeconomic background may affect the placement of biracial individuals within a racial hierarchy. Socioeconomic background plays a role in racial inequality more generally as we know that accounting for socioeconomic difference often reduces the magnitude of observed racial inequality, suggesting that some of racial inequality we observe operates through class differences (CITATION). This issue is particularly important for biracial individuals because of potential class selection into interracial unions. The likelihood of being an interracial union varies by class background for at least some groups (CITATION), although it varies in complex ways and we know more about interracial marriages than other types of unions that may produce children. For example, black partners tend to be positively selected with regard to education into Black/White marriages while white partners are negatively selected (CITATION). According to status exchange theory (CITATIONS), this pattern may be common to all types of interracial marriages within an unequal status hierarchy, although the evidence of status exchange for other forms of marriages is weaker (CITATION). How this affects the relative resources of biracial individuals will depend then in some part on the relative strength of these phenomena for a given pairing, the gender balance in that pairing, and how the gender of parents affects outcomes.

Regardless of the underlying complexity, delineating the role of this selection process in the prior generation as opposed to other mechanisms is key to understanding the overall process that sorts biracial individuals into a racialized system. Although this approach is sometimes characterized as pitting race vs class and minimizing the importance of race, that is not case here. Instead, what this delineation allows us to do is understanding how race affects the outcomes of biracial individuals in a multigenerational process and allow us to better pinpoint the generational locus of certain mechanisms.

## Prior Work

@@

## Grade Retention as an Outcome
